Sharing statistics for SPARQL Federation optimization, with emphasis on benchmark quality

### 2012-01-12 eswc-2012/phdsymp.tex
Benchmark statistics design
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@72 43ef1190-90fd-4dd7-a769-d0121f02b80e

Using techniques from statistical experimental design, I conjecture
that it is possible to design an experiment (i.e. a benchmark) so that
we can prove it covers all realistic cases and that we can justify
why the remaining corner cases are unlikely to influence the result.

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of benchmarking.

@@

\bibliographystyle{plain}
\bibliography{selectivity,federation,benchmarks}

>>>>

Using techniques from statistical experimental design, I propose
that it is possible to design an experiment (i.e. a benchmark) so that
we can prove it covers all realistic cases and that we can justify
why the remaining corner cases are unlikely to influence the result.

@@

This part of the thesis seeks to use advances from statistical
experimental planning to improve accuracy and dependability of
benchmarking. Already in 1926, Ronald~Fischer noted that complex
experiments can be much more efficient than simple
ones~\cite{Fisher_1926}\footnote{Cited in
  http://en.wikipedia.org/wiki/Factorial\_experiment}, starting the
experimental design field. One of the simpler designs is ``fractional
factorial design'', in which several ``factors'' are studied by using
different ``treatments''. In terms of SPARQL execution, one factor can
for example be the level of nested \texttt{OPTIONAL}s, number of
triples in a basic graph pattern, etc, and the treatment is to vary
these numbers. The key to understanding why this can be efficient is
that these variations need not occur in the same experiment. Thus,
many combinations of factors can be studied by carefully designing
queries to cover different factors, and a formalism called
``resolution'' has been developed to classify how well this has been
achieved.

Fractional factorial design is covered in elementary text-books in
statistics, and I intend to go further into the statistical literature
to see if there is methodology that is even better suited to the
problem. Armed with this, I hope to use amongst other things, the
complexity analysis of SPARQL (see
e.g. \cite{Schmidt:2010:FSQ:1804669.1804675}) to find suitable factors
to see if my admittedly bold proposition that it is possible to design
a benchmark to cover all realistic cases.

@@

\bibliographystyle{plain}
\bibliography{selectivity,federation,benchmarks,statistics,optimization}
